Constraints on the mass and occupation fraction of dark matter halos hosting Lyman-alpha emitters at $z\sim 3$

### 2013-07-24 paper/ms.tex
ACF calculation described, paulina's comments added

\title[Halo mass and occupation fraction for LAEs at
  $z=3.1$]{Constraints on the mass and occupation
  fraction of dark matter halos hosting Lyman-alpha emitters
  at $z\sim 3$}    
\author[~J.~E. Forero-Romero and J. Mejia]{
\parbox[t]{\textwidth}{\raggedright 
Jaime E. Forero-Romero$^{1}$ and
Julian Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

Lyman-$\alpha$ emitting galaxies (LAEs) have become in the last decade
a  central topic in studies of structure formation in the Universe. They 
are helpful in a diverse range of fields. LAEs can be used as probes
of reionization \citep{Dijkstra11}, tracers of large scale structure
\citep{Koehler2007},  signposts for low metallicity stellar
populations and markers of thethe galaxy formation process through
cosmic history \citep{ForeroRomero2012}.

At the same time, theoretical and observational developments have
contributed to the emergence of a paradigm to describe structure
formation in a cosmological context. In this context it is considered
that dominant matter content of the Universe is to be found in dark
matter, whereby each galaxy is hosted by larger dark matter structure
known as a halo.

@@

The second constraint is the angular correlation function (ACF). The
measurements presented in \citep{Yamada2012} do not report an ACF
measurement. Instead we use the results reported by
\citep{Hayashino2004} on the LAEs observed in the densest field of
SSA22. It is important to keep in mind that there are some differences
between this work and \citep{Yamada2012}. First, the color selection by
\cite{Yamada2012} is less stringent compared to the one by
\cite{Hayashino2004}. Also the EW threshold is different,
\cite{Hayashino2004} uses a cut around $154$\AA \ instead of $190$\AA
\ used by \citep{Yamada2012}

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking length of 0.17 times the inter-particle
distance. The minimum halo mass in the models we construct in this \documentname
correspond to groups of $\sim 75$ particles. The catalogs were
obtained from the publicly available Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.
We focus our work on halos more massive than $1\times 10^{10}$\hMsun
that are resolved with at least $70$ particles (REFF????)

@@

The comparison between the theoretical and observational ACF is done
in therms of the angular correlation length obtained  by fitting to a power-law
function:

@@

Both the theoretical and observational ACF are derived by a least
square minimization procedure. We consider that a model is consistent
with observations if the two parameters $\gamma$ and $\theta_0$ are
equal within a $1$-$\sigma$ range.

@@

\begin{figure}
\begin{center}
\includegraphics[width=1.10\linewidth,angle=0]{./plots/Fig1.pdf}
\caption{ \label{fig:halos} Surface density of dark
  matter halos as a function of a minimum halo mass to count the
  total number of elements in a volume. Each line represents on of the
  $210$ volumes of dimensions $46\times 35\times 41$ h${-3}$Mpc$^{3}$
  in the Bolshoi simulation. The horizontal gray band represents the
  range of surface densities observed for LAEs at $z=3.1$ as reported
  by \citep{Yamada2012}.}
\end{center} 
\end{figure}

@@

In Figure \ref{fig:restriction_mock_and_f_occ_corr} we present the preferred
models in the planes $M_{\rm min}-M_{\rm  max}$ and $M_{\rm min}-f_{\rm occ}$
for the {\tt match} method after applying the observational constraints 
in the occupation fraction and the ACF. Here we consider a model
consistent if there is an overlap of $1-\sigma$ in its values for the
correlation length $\theta_0$ and the power $\gamma$ in the
correlation function.

@@

We also find that $70\%$ of the best models are found in families (1)
and (2), with a low occupation fraction $f_{\rm occ}\leq 0.3$. This
preference goes in the same direction as the observational constraint
on $f_{\rm occ}\sim 0.1-0.2$ derived at $z=2.2$ by \cite{Hayes2010}
and recent theoretical study of observational data in a wide redshift
range $0<z<6$  \citep{Dijkstra2013}.

@@

We also find interesting that our ACF analysis is also not enough to rule
out models with a high occupation fraction $f_{\rm occ}>0.3$, which
represent almost one fourth of our best models, coinciding with a wide
range in halo masses $\Delta M>1.0$ dex. These models can only be
considered unfavorable based on a different set of observations as we
have described in the previous subsection.

@@

Considering the additional evidence for a low escape fraction we can
say that the preferred models are in families (1) and (2) with a
clear majority composed by those with narrow mass range.

@@

A central result in this exploration of parameter esp ace is the
existence of a large number of models with a very
narrow mass range $\Delta M< 0.1$ dex that are consistent in every
aspect with the spatial distribution of observed LAEs. In order to
have these characteristics it is required that halos around the mass
$M_{\rm max}$ for each model become inefficient in hosting detectable
galaxies. One way to achieve this is having LAEs with a decreasing \ly
escape fraction with increasing mass.

>>>>

\title[Halo mass and occupation fraction for LAEs at
  $z=3.1$]{Constraints on the mass and occupation
  fraction of dark matter halos hosting Lyman-alpha emitters
  at $z\sim 3$}    
\author[~J.~E. Forero-Romero and ~J.~E Mejia-Restreo]{
\parbox[t]{\textwidth}{\raggedright 
Jaime E. Forero-Romero$^{1}$ and
Julian E. Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

Lyman-$\alpha$ emitting galaxies (LAEs) have become in the last decade
a  central topic in studies of structure formation in the Universe. They 
are helpful in a diverse range of fields. LAEs can be used as probes
of reionization \citep{Dijkstra11}, tracers of large scale structure
\citep{Koehler2007},  signposts for low metallicity stellar
populations, markers of the galaxy formation process through
cosmic history \citep{ForeroRomero2012}. and tracers of active star formation 
\citep{Guaita2013}

At the same time, theoretical and observational developments have
contributed to the emergence of a paradigm to describe structure
formation in a cosmological context. In this context it is considered
that dominant matter content of the Universe is to be found in dark
matter, whereby each galaxy is hosted by larger dark matter structure
known as a halo. \citep{Peebles1980}

@@

The second constraint is the angular correlation function (ACF). The
measurements presented in \citep{Yamada2012} do not report an ACF
measurement. Instead we use the results reported by
\citep{Hayashino2004} on the LAEs observed in the densest field of
SSA22. It is important to keep in mind that there are some differences
between this work and \citep{Yamada2012}. First, the color selection by
\cite{Yamada2012} is less stringent compared to the one by
\cite{Hayashino2004}. Also the EW threshold is different,
\cite{Hayashino2004} uses a cut around $154$\AA \ instead of $190$\AA
\ used by \citep{Yamada2012}. These differences  translates
in only two more detected objects in the Hayashino survey (283 LAEs) than in 
Yamada et al (281 LAEs).

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking length of 0.17 times the inter-particle
distance. The minimum halo mass in the models we construct in this \documentname
correspond to groups of $\sim 75$ particles. The catalogs were
obtained from the publicly available Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.
We focus our work on halos more massive than $1\times 10^{10}$\hMsun
that are resolved with at least $70$ particles ({\bf REFF???? or support???})

@@

We then calculate the  angular correlation function (ACF)
for all the models having the $15$ mock surveys consistent with
observations . The ACF is computed  ( using 
the Davis \& Peebles  estimator \citep{Davis1983} ) only on the densest
subfield in all the 15 mock surveys corresponding to the SSA22
region. We then take the mean ACF over the 15  mock surveys  for
each model as well as its the standard deviation to include
the effect of cosmic variance.

We finally compare the observational ACF and the mean mock  ACF of the models   
in therms of the angular correlation lenght ($\theta_{0}$) by fitting 
them to a power-law function:

@@

Both mock and observational ACF are derived by a least
square minimization procedure. The observational errors of 
the ACF as well as the standard deviation of the mean modelled ACFs  are
considered to compute the error in the fitted parameters to ensure ourselves
to include poissionian and cosmic-variance errors. 
We consider that a model is consistent with observations if the two parameters $\gamma$ and $\theta_0$ are
equal within a $1$-$\sigma$ range.

@@

\begin{figure}
\begin{center}
\includegraphics[width=1.10\linewidth,angle=0]{./plots/Fig1.pdf}
\caption{ \label{fig:halos} Surface density of dark
  matter halos as a function of a minimum halo mass to count the
  total number of elements in a volume. Each line represents one of the
  $210$ volumes of dimensions $46\times 35\times 41$ h${-3}$Mpc$^{3}$
  in the Bolshoi simulation. The horizontal gray band represents the
  range of surface densities observed for LAEs at $z=3.1$ as reported
  by \citep{Yamada2012}.}
\end{center} 
\end{figure}

@@

In Figure \ref{fig:restriction_mock_and_f_occ_corr} we present the preferred
models in the planes $M_{\rm min}-M_{\rm  max}$ and $M_{\rm min}-f_{\rm occ}$
for the {\tt match} method after applying the observational constraints 
in the occupation fraction and the ACF. Here we consider a model
consistent if there is an overlap of $1-\sigma$ in its values for the
correlation length $\theta_0$ and the power $\gamma$ in the
correlation function.  As before, we found similar results using the {\tt random} method

@@

We also find that $70\%$ of the best models are found in families (1)
and (2), with a low occupation fraction $f_{\rm occ}\leq 0.3$. This
preference goes in the same direction as the observational constraint
on the escape fraction $f_{\rm esc}\sim 0.1-0.2$ derived at $z=2.2$ by \cite{Hayes2010}
and recent theoretical study of observational data in a wide redshift
range $0<z<6$  \citep{Dijkstra2013}.

@@

We also find interesting that our ACF analysis is also not enough to rule
out models with a high occupation fraction $f_{\rm occ}>0.3$, which
represent almost one fourth of our best models, coinciding with a wide
range in halo masses $\Delta M>1.0$ dex. These models can only be
considered unfavorable based on the density and abundance observations
as we have described in the previous subsection.

@@

Considering the additional evidence for a low escape fraction and 
assuming a direct relation between $f_{esc}$ and $f_{occ}$ we can
say that the preferred models are in families (1) and (2) with a
clear majority composed by those with narrow mass range.

@@

A central result in this exploration of parameter espace is the
existence of a large number of models with a very
narrow mass range $\Delta M< 0.1$ dex that are consistent in every
aspect with the spatial distribution of observed LAEs. In order to
have these characteristics it is required that halos around the mass
$M_{\rm max}$ for each model become inefficient in hosting detectable
galaxies. One way to achieve this is having LAEs with a decreasing \ly
escape fraction with increasing mass.
